---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

## Loading and preprocessing the data

Show any code that is needed to

  1.  Load the data (i.e. `read.csv()`)

  2.  Process/transform the data (if necessary) into a format suitable for your analysis

```{r}
activity <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in the dataset.

  1.  Calculate the total number of steps taken per day

```{r}
count <- 0
steps <- c()
for (i in levels(activity$date)) {
  stepstoday <- subset(activity, date==i)$steps
  count = count + 1
  steps[count] <- sum(stepstoday[!is.na(stepstoday)])
}
steps4days <- data.frame(levels(activity$date), steps)
summary(steps4days)
```

  2.  If you do not understand the difference between a histogram and a barplot, research the
  difference between them. Make a histogram of the total number of steps taken each day

```{r}
hist(steps)
```

3.  Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean.steps <- mean(steps4days$steps)
mean.steps
median.steps <- median(steps4days$steps)
median.steps
```

## What is the average daily activity pattern?

  1.  Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the
  average number of steps taken, averaged across all days (y-axis)

```{r}
daily <- data.frame()
for (i in levels(activity$date)) {
  stepstoday <- subset(activity, date==i)$steps
  daily <- rbind(daily, stepstoday)
}
intervalnum <- 0
average <- c()
for (i in 1:dim(daily)[2]){
  intervalnum = intervalnum + 1
  thisinterval <- daily[,i]
  meanthisone <- mean(thisinterval[!is.na(thisinterval)])
  average <- c(average, meanthisone)
}
plot(ts(average))
```

  2.  Which 5-minute interval, on average across all the days in the dataset, contains the
  maximum number of steps?

```{r}
max(average)
which.max(average)
```

## Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). 
The presence of missing days may introduce bias into some calculations or summaries of the data.

  1.  Calculate and report the total number of missing values in the dataset (i.e. the total
    number of rows with NAs)

```{r}
miss <- sum(is.na(activity$steps))
miss
```

  2.  Devise a strategy for filling in all of the missing values in the dataset. The strategy 
    does not need to be sophisticated. For example, you could use the mean/median for that
    day, or the mean for that 5-minute interval, etc.

We decided to fill de missing value with de average of that interval across all days.

  3.  Create a new dataset that is equal to the original dataset but with the missing data
    filled in.

```{r}
av <- data.frame(activity[1:288, 3], average)
nomiss <- activity
for (i in which(is.na(activity$steps))){
  nomiss[i,1] <- av[match(nomiss[i,3], table=av[,1]),2]
}
```

  4.  Make a histogram of the total number of steps taken each day and Calculate and report
    the mean and median total number of steps taken per day. Do these values differ from
    the estimates from the first part of the assignment? What is the impact of imputing 
    missing data on the estimates of the total daily number of steps?

Number of steps per-day:

```{r}
count <- 0
steps.nomiss <- c()
for (i in levels(nomiss$date)) {
  stepstoday <- subset(activity, date==i)$steps
  count = count + 1
  steps.nomiss[count] <- sum(stepstoday[!is.na(stepstoday)])
}
steps4days.nomiss <- data.frame(levels(nomiss$date), steps.nomiss)
summary(steps4days.nomiss)
```

Mean and median from all days:

```{r}
mean(steps4days.nomiss$steps.nomiss)
median(steps4days.nomiss$steps.nomiss)
```

Histogram:

```{r}
hist(steps.nomiss)
```

5-minutes interval:

```{r}
daily.nomiss <- data.frame()
for (i in levels(nomiss$date)) {
  stepstoday <- subset(nomiss, date==i)$steps
  daily.nomiss <- rbind(daily.nomiss, stepstoday)
}
intervalnum <- 0
average.nomiss <- c()
for (i in 1:dim(daily.nomiss)[2]){
  intervalnum = intervalnum + 1
  thisinterval <- daily.nomiss[,i]
  meanthisone <- mean(thisinterval[!is.na(thisinterval)])
  average.nomiss <- c(average.nomiss, meanthisone)
}
plot(ts(average.nomiss))
```

Analysis of the impact of missing values: There is no difference in the variables

Mean and median from all days:

```{r}
mean(steps4days.nomiss$steps.nomiss) - mean(steps4days$steps)
median(steps4days.nomiss$steps.nomiss) - median(steps4days$steps)
```

5-minutes interval:

```{r}
plot(ts(average.nomiss - average))
```

Steps for days:

```{r}
sum(steps4days.nomiss$steps.nomiss - steps4days$steps)
sum(steps4days.nomiss$steps.nomiss - steps4days$steps)
```
There are no differences at all.

## Are there differences in activity patterns between weekdays and weekends?

For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing
values for this part.

  1.  Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating
  whether a given date is a weekday or weekend day.

```{r}
count = 0
weekend <- c()
for (i in weekdays(as.POSIXct.Date(activity$date))){
  count = count + 1
  if (i == "sábado" | i == "domingo"){
    weekend[count] <- "weekend"
  } else {
    weekend[count] <- "weekday"
  }
}
nomiss <- cbind(nomiss, weekend)
```

  2.   Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis)
  nd the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the
  README file in the GitHub repository to see an example of what this plot should look like using simulated
  data.
  

```{r}
weekdays <- subset(nomiss, weekend=="weekday")
daily.weekdays <- data.frame()
for (i in levels(weekdays$date)) {
  stepstoday <- subset(weekdays, date==i)$steps
  daily.weekdays <- rbind(daily.weekdays, stepstoday)
}
intervalnum <- 0
average.weekdays <- c()
for (i in 1:dim(daily.weekdays)[2]){
  intervalnum = intervalnum + 1
  thisinterval <- daily.weekdays[,i]
  meanthisone <- mean(thisinterval[!is.na(thisinterval)])
  average.weekdays <- c(average.weekdays, meanthisone)
}

weekends <- subset(nomiss, weekend=="weekend")
daily.weekends <- data.frame()
for (i in levels(weekends$date)) {
  stepstoday <- subset(weekends, date==i)$steps
  daily.weekends <- rbind(daily.weekends, stepstoday)
}
intervalnum <- 0
average.weekends <- c()
for (i in 1:dim(daily.weekends)[2]){
  intervalnum = intervalnum + 1
  thisinterval <- daily.weekends[,i]
  meanthisone <- mean(thisinterval[!is.na(thisinterval)])
  average.weekends <- c(average.weekends, meanthisone)
}
par(mfrow=c(2,1)) 
plot(ts(average.weekdays))
plot(ts(average.weekends))
```

